Spike spectra for recurrences

### 2022-10-26 Manuscript/1st resubmission/rebuttal.md

_Chapter 2 presents the signal decomposition into set of appropriate basis functions using Dirac comb. Here, too, we can distinguish a graphic representation of the decomposition procedure (Fig. 3). Nevertheless, the Authors could cite a position in the literature that discusses other methods of decomposition._

TBD

_A small note on Appendix C: It would be interesting to present an inter-spike spectra also for one of the nonlinear systems from Duffing, Mackey-Glass or Chen systems._

We thank the reviewer for this suggestion. In fact, we were computing numerous inter-spike spectra for different systems. But regarding the manuscript, we decided to present a variety of possible applications, which includes the analysis of exemplary nonlinear system (like the Roessler we showed or systems you suggest) as well as "real-world" data (power grid frequency data) and data stemming from a geophysical context (orbital data). Each of these applications could, of course, be expanded or other applications could be shown, but here we must concentrate on presenting the method itself as concisely as possible while keeping the size of the manuscript within tolerable limits. Moreover, when analyzing a nonlinear system we would potentially be interested in different dynamical setups, which would, in turn, lead to a variety of aspects we would additionally need to consider and cover with the proposed method. The idea of this paper is to outline the proposed method and provide the code/package for its use. We would certainly be pleased if researchers decided to publish a thorough study of a number of different nonlinear systems using this method.

### 2022-10-27 Manuscript/1st resubmission/rebuttal.md

_Introduction: a clear formulation of the mathematical problem, the state-of-the-art in terms of the existing literature, a summary of the solution proposed and of its pros and cons are totally lacking._

We thank for this comment, but do not agree completely with this criticism. In our opinion, the problem is well stated and the outline of the manuscript is summarised. But the literature review can be improved, which with the help of the suggestions of the reviewer 2 we finally did. "a summary of the solution proposed" as well as "pros and cons" do not belong to the introduction. These are provided in the conclusion of the manuscript. We added some further introductory sentences to the beginning of the introduction.

_Equation (1) is poorly introduced. Although the authors cite a reference ([1]), the x_i, x_j vectors, as well as the parameters d and N, are not explained. The "trajectory x_i", for example, is introduced only at the end of page 1. Moreover, R_{i,j}(\epsilon) is not explicitly stated and explained._

Indeed, out intention was to have a short but clear introduction into the recurrence plot method, which reviewer 2 obviously liked very much. Nevertheless, we have now added an explanation on the recurrence plot definition. An unfamiliar reader should now better understand the method.

@@

We have extended the introduction and hope it is now more clear.

_Moreover, the Authors explained the algorithm in great detail (in numerous pictures). By the way, also with outer -spike. Literature examples below review._

We have added some further references on spike-train analysis and, in particular, on spike-train power spectra. Thanks for the literature suggestions, which were really helpful!

_The illustrations of tau-recurrence rate based spectrum (Fig. 1) and the transformation diagrams of the series of Dirac delta function (Fig. 2), including the two proposed inter-spikes spectrum, deserve emphasis. A small remark - regarding verse 35: what will be with stochastic resonance, because we are dealing with impulse disturbances.?_

### 2022-10-27 Manuscript/inter_spike_manuscript_mdpi.tex
More refs for answering reviewer 2 comments

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{In reccurance analysis, the $\tau$-recurrence rate encodes the periods of the cycles of the underlying high-dimensional timeseries. It thus plays a similar role to the autocorrelation for scalar time-series in encoding temporal correlations. However, its Fourier decomposition does not have a clean interpretation. Thus there is no satisfactory analogue to the power spectrum in recurrence analysis. In this paper we introduce a novel way to decompose the $\tau$-recurrence rate using an overcomplete basis of Dirac combs together with a sparsity regularization. We show that this decomposition, the \textit{inter-spike spectrum}, naturally provides an analogue to the power spectrum for recurrence analysis in the sense that it reveals the dominant periodicities of the underlying time series. We show that the inter-spike spectrum correctly identifies patterns and transitions in the underlying system in a wide variety of examples and is robust to measurement noise.}
%
%A novel kind of power spectrum is constructed, the \textit{inter-spike spectrum}, which transforms any signal into its spike-frequency domain. This method 
%clearly shows the apparent cycles in the data and overcomes the problems for spike-train-like signals when using the obvious idea of Fourier-transforming it. 
%We invent this instructive approach with the 
%idea of transforming the $\tau$-recurrence rate of a recurrence plot (RP), which often has a spiky appearance. The $\tau$-recurrence rate is the density of recurrence points 
%along diagonals of the RP, which are parallel to the main diagonal with a distance of $\tau$. In this context the inter-spike spectrum can be interpreted as a nonlinear power 
%spectrum of a potentially high dimensional system which constitutes the RP. The proposed measure is robust to noise and is able to detect and analyze bifurcations.}

@@

For example, consider the signal we would like to analyze 
(e.g., the $\tau$-RR of a system) to be a Dirac comb (DC) with inter-spike period $T_\text{is}$: 
\begin{equation}
\text{DC}_{\text{is}}(t) = \sum_{k=-\infty}^{\infty} \delta(t-kT_\text{is}),
\label{eq_dirac_comb}
\end{equation}
i.e., a series of Dirac delta functions for a period $T_\text{is}$. There is only one single period, $T_\text{is}$, in this signal (Fig.~\ref{fig_tau_rr_dirac_comb}A, D), so 
in principle we would strive for a single peak in the frequency domain of this signal at a frequency $f=1/T_\text{is}$. The Fourier spectrum does not meet this expectation 
and instead of a single frequency, there are many frequencies excited (Fig.~\ref{fig_tau_rr_dirac_comb}B, E). This is because the Fourier components add constructively 
for every frequency $1/T_\text{is}$ and therefore $\text{DC}_{\text{is}}(t)$ coincides with its own Fourier transform up to a factor $1/T_\text{is}$.

>>>>

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{In recurrence analysis, the $\tau$-recurrence rate encodes the periods of the cycles of the underlying high-dimensional time series. It thus plays a similar role to the autocorrelation for scalar time-series in encoding temporal correlations. However, its Fourier decomposition does not have a clean interpretation. Thus there is no satisfactory analogue to the power spectrum in recurrence analysis. In this paper we introduce a novel way to decompose the $\tau$-recurrence rate using an over-complete basis of Dirac combs together with a sparsity regularization. We show that this decomposition, the \textit{inter-spike spectrum}, naturally provides an analogue to the power spectrum for recurrence analysis in the sense that it reveals the dominant periodicities of the underlying time series. We show that the inter-spike spectrum correctly identifies patterns and transitions in the underlying system in a wide variety of examples and is robust to measurement noise.}
%
%A novel kind of power spectrum is constructed, the \textit{inter-spike spectrum}, which transforms any signal into its spike-frequency domain. This method 
%clearly shows the apparent cycles in the data and overcomes the problems for spike-train-like signals when using the obvious idea of Fourier-transforming it. 
%We invent this instructive approach with the 
%idea of transforming the $\tau$-recurrence rate of a recurrence plot (RP), which often has a spiky appearance. The $\tau$-recurrence rate is the density of recurrence points 
%along diagonals of the RP, which are parallel to the main diagonal with a distance of $\tau$. In this context the inter-spike spectrum can be interpreted as a nonlinear power 
%spectrum of a potentially high dimensional system which constitutes the RP. The proposed measure is robust to noise and is able to detect and analyze bifurcations.}

@@

For example, consider the signal we would like to analyze 
(e.g., the $\tau$-RR of a system) to be a Dirac comb (DC) with inter-spike period $T_\text{is}$: 
\begin{equation}
\text{DC}_{\text{is}}(t) = \sum_{k=-\infty}^{\infty} \delta(t-kT_\text{is}),
\label{eq_dirac_comb}
\end{equation}
i.e., a series of Dirac delta functions for a period $T_\text{is}$. There is only one single period, $T_\text{is}$, in this signal (Fig.~\ref{fig_tau_rr_dirac_comb}A, D), so 
in principle we would strive for a single peak in the frequency domain of this signal at a frequency $f=1/T_\text{is}$. The Fourier spectrum does not meet this expectation 
and instead of a single frequency, there are many frequencies excited (Fig.~\ref{fig_tau_rr_dirac_comb}B, E). This is because the Fourier components add constructively 
for every frequency $1/T_\text{is}$ and therefore $\text{DC}_{\text{is}}(t)$ coincides with its own Fourier transform up to a factor $1/T_\text{is}$.\footnote{This can also 
be observed for neuron spike trains, e.g. \cite{Orcioni2020,Biagetti2017}.}
